\bibstyle{amsalpha}

\chapter{Dense linear orders}

The prototypical dense linear orders are the rationals $\eta$ and the reals
$\lambda$.  The former is the unique (up to isomorphism) countable dense linear
order, whereas the latter is its Dedekind completion.

In this chapter we discuss the model theory surrounding the elementary class of
dense linear order orders.  We prove an analogue of the theorem of La\"uchli and
Leonard for this class by extending the class $\Mzero$ using an operation known
as the shuffle operation.  This lays the necessary ground work in order to prove
that the theory of \textit{all} linear orders is decidable.

\section{The shuffle operation}

\begin{prp}\label{prp:ratpart}
	Suppose $n\geq 2$ is a natural number and let $P\subseteq\nats$ bet the set
	of the first $n-1$ primes $p_0,\dotsc,p_{n-2}$.  Choose, for each $i<n-1$,
	the set
	\begin{equation}
		Q_i=\bigcup_{\substack{k\in\ints\\ k\neq
				0}}\setbuild{\faktor{a}{p_i^k}}{a\in\ints\text{ and } p_i\nmid a}.
	\end{equation}
	and define $Q_{n-1}=\rats\setminus\bigcup_{i<n-1}Q_i$.  Then
	$\mathcal{Q}=\setbuild{Q_i}{0\leq i<n}$ is a partition of $\rats$ into dense
	subsets of $\eta$.
\end{prp}
\begin{proof}
	By definition, it clearly follows that $\mathcal{Q}$ is a partition of
	$\rats$.  Hence, all that remains is to show that its equivalence classes
	are dense subset of the rationals.

	We commence by showing that $Q_{n-1}$ is such a dense subset.  Choose $p$ to
	be the least prime such that $p>p_{n-1}$ and define
	\begin{equation}
		S=\bigcup_{\substack{k\in\ints\\ k\neq
				0}}\setbuild{\faktor{a}{p^k}}{a\in\ints\text{ and } p\nmid a}.
	\end{equation}
	Since $S\subseteq Q_{n-1}$, it is sufficient to prove that $S$ is a dense
	subset of $\eta$.  If $s_0<s_1$ are elements of $S$ then there exists
	$a_0,a_1\in\ints$ and $k_0,k_1\in\posnats$ such that, for $i=0,1$, it holds
	that $p\nmid a$ and
	\begin{equation}
		s_i=\frac{a_i}{p^{k_i}}.
	\end{equation}

	Suppose first that $\abs{s_1-s_0}\geq 1$ and choose $m>k_0$ to be the least
	natural number such that $p^m>\abs{s_1-s_0}$.  Define
	\begin{equation}
		t=s_0+\frac{1}{p^m}.\label{eq:qpart}
	\end{equation}
	Then it follows that $s_0<t<s_1$ and
	\begin{equation}
		t=\frac{p^{m-k_0}a_0+1}{p^m}.
	\end{equation}
	Observe that, if $p$ divides the numerator, then we may conclude that $p\mid
		1$.  This is clearly a contradiction and, thus, $p$ cannot divide the
	aforementioned numerator.

	Suppose now, instead, that $\abs{s_1-s_0}<1$ and choose $m>k_0$ to be the
	least natural number such that $p^{-m}<\abs{s_1-s_0}$.  Define $t$ as
	in (\ref{eq:qpart}) then, similar to previously, it follows that
	\begin{equation}
		t=\frac{p^{m-k_0}a_0+1}{p^m}.
	\end{equation}
	Again, using a similar argument, we may conclude that $p$ does not divide
	the numerator above.

	An analagous methodology may be employed in order to prove that $Q_i$ is a
	dense subset of $\eta$ for $i<n$.  One need only substitute each $Q_i$ for
	$S$, and $p_i$ for $p$, in the above proof.
\end{proof}

As we will be revisiting it, we shall refer to the partition of $\rats$, defined
in the proposition above, as the \textit{canonical partition} of $\rats$ into
$n$ dense subsets.  Similarly, the \textit{canonical partition of the reals}
into $n$ dense subsets is laid out in the following corollary.

\begin{cor}
	For every natural $n\geq 2$, there exists a canonical partition
	$\mathcal{R}=\set{R_0,\dotsc,R_{n-1}}$ of $\reals$ such that:
	\begin{enumerate}
		\item $R_{n-1}=\irrats$,
		\item $R_k$ is a countable dense subset of $\lambda$, for each natural $k<n-1$.
	\end{enumerate}
\end{cor}
\begin{proof}
	Choose $R_0,\dotsc,R_{n-2}$ so that they correspond exactly to the
	equivalence classes of the canonical partition of $\rats$ into dense subsets
	of $\eta$.  Together with $R_{n-1}=\irrats$, these then form the equivalence
	classes of the desired partition $\mathcal{R}$ of $\reals$.
\end{proof}

\begin{dfn}[The shuffle operation]
	Let $n\in\nats$ and choose $F$ to be any finite set of linear orders.  Now
	define the \textbf{shuffle operation} on the class of linear orders as
	follows:
	\begin{enumerate}
		\item if $F=\emptyset$ then $\sigma(F)=\zero$;
		\item if $F=\set{\alpha}$ then $\sigma(F)=\alpha\cdot\eta$;
		\item if $n\geq 2$ and $h\colon\rats\to F$ is any colouring of $\eta$
		      whose induced partition on $\rats$ is the canonical one, define
		      \begin{equation}
			      \sigma(F)=\sum_{q\in\eta}h(q).
		      \end{equation}
	\end{enumerate}
	Furthermore, the image of $F$ under $\sigma$ is referred to as the
	\textbf{shuffle} of $F$.
\end{dfn}


\section{The L\"auchli and Leonard result for the class of linear orders}

\begin{dfn}[The class $\M$]
	The class $\M$ is the smallest class of linear orders such that the following holds:
	\begin{enumerate}
		\item	$\zero,\one\in\M$,
		\item	if $\alpha,\beta\in\M$ then $\alpha+\beta\in\M$,
		\item	if $\alpha\in\M$ then $\alpha\cdot\omega,\alpha\cdot\dual{\omega}\in\M$,
		\item	for every finite $F\subseteq\M\setminus\set{\zero}$, it holds that $\sigma(F)\in\M$.
	\end{enumerate}
\end{dfn}

\begin{prp}\label{prp:Msumint}
	If $\alpha_i\in\M$ for each $i\in\ints$ then there exists a $\beta\in\M$ such that
	\begin{equation}
		\sum_{i\in\zeta}\alpha_i\nequiv{n}\beta.
	\end{equation}
\end{prp}
\begin{proof}
	Since $\M$ is closed under finite sums as well as right multiplication by $\omega$ and $\dual{\omega}$, results analagous to propositions \ref{prp:M0sum} and \ref{prp:M0sumop} can be proven in the context of $\M$.  An argument similar to that which appears in the proof of proposition \ref{prp:M0sumint} then yields the desired result.
\end{proof}

\begin{thm}[L\"auchli and Leonard]\label{thm:LLlin}
	For every countable linear order $\alpha$, and every $n\in\nats$, there exists some $\beta_n\in\M$ such that $\alpha\nequiv{n}\beta_n$.
\end{thm}
\begin{proof}
	Define a binary relation $R$ on $\alpha$ such that, for every $x,y\in\alpha$, $xRy$ iff $x\leq y$ and there exists a $\beta\in\M$ such that $[x,y]\nequiv{n}\beta$.  Since $R$ is clearly transitive (simply note $\M$ is closed under finite sums), it induces a congruence $\sim$ on $\alpha$.    We now claim the following:
	\begin{claim}
		For every $\gamma\in\faktor{\alpha}{\sim}$, there exists a $\beta\in\M$ such that $\gamma\nequiv{n}\beta$.
	\end{claim}
	\begin{proof}
		Fix any $\gamma\in\faktor{\alpha}{\sim}$ and note that $\M$ contains all finite ordinals so we may assume, without loss of generality, that $\gamma$ is infinite.  Consequently, at least one of $\omega$ and $\dual{\omega}$ is embeddable in $\gamma$.  There must then exist a family $\family{\gamma_i}{i\in\zeta}$ of (possibly empty) linear orders such that
		\begin{equation}
			\gamma\cong\sum_{i\in\zeta}\gamma_i.
		\end{equation}
		By definition of $\sim$ it now follows that there exists a $\gamma_i^\prime\in\M$, for each $i\in\ints$, such that $\gamma_i\nequiv{n}\gamma_i^\prime$.  Proposition \ref{prp:Msumint} then implies that there exists a $\gamma^\prime\in\M$ such that
		\begin{equation}
			\gamma^\prime\nequiv{n}\sum_{i\in\zeta}\gamma_i^\prime.
		\end{equation}
		From lemma \ref{lem:fvsum} we may conclude that
		\begin{equation}
			\gamma\nequiv{n}\sum_{i\in\zeta}\gamma_i^\prime\nequiv{n}\gamma^\prime,
		\end{equation}
		establishing the claim.
	\end{proof}

	Note that if $\card{\faktor{\alpha}{\sim}}=1$ then our task is complete so suppose, by way of contradiction, that $\card{\faktor{\alpha}{\sim}}>1$.
	\begin{claim}
		$\faktor{\alpha}{\sim}$ is dense.
	\end{claim}
	\begin{proof}
		With our goal being a contradiction, suppose that $\faktor{\alpha}{\sim}$ is not dense.  Therefore, there exists $a,b\in\alpha$ such that $a<b$, $a\not\sim b$ and for every $c\in(a,b)$ either $c\sim a$ or $c\sim b$ but not both.  Note that, since $a\not\sim b$, the interval $(a,b)$ cannot be finite.  Therefore, $(a,b)\cong\sum_{i\in\zeta}\gamma_i$ where $\gamma_i$ is $n$-equivalent to some member $\gamma_i^\prime$ of $\M$, for each $i\in\ints$.  Thus, by proposition \ref{prp:Msumint}, $(a,b)\nequiv{n}\gamma$ for some $\gamma\in\M$.  Since $\one+\gamma+\one\nequiv{n}[a,b]$ by lemma \ref{lem:fvsum} and $\one+\gamma+\one$ is a member of $\M$, it follows by definition that $a\sim b$ --- a contradiction.
	\end{proof}

	Now choose $K=\set{\chi_0,\dotsc,\chi_{m-1}}$, for some $m\in\posnats$, to be an $n$-spectrum for the class $\setbuild{\chi\in\M}{\chi\nequiv{n}\gamma,\text{ for some }\gamma\in\faktor{\alpha}{\sim}}$.
	\begin{claim}
		There exists a maximal $F\subseteq K$ and a nonempty interval $I=(A,B)$ of $\faktor{\alpha}{\sim}$ such that, for every $\chi\in F$, the set $D_\chi(I)=\setbuild{\gamma\in I}{\gamma\nequiv{n}\chi}$ is dense in $I$.
	\end{claim}
	\begin{proof}
		Argue by induction on $m$.  If $m=1$ then, since $\faktor{\alpha}{\sim}$ is dense we can choose any $A,B\in\faktor{\alpha}{\sim}$ such that $A<B$ and it will immediately follow that if $I=(A,B)$ then $D_{\chi_0}(I)$ is dense in $I$, since every member of $\faktor{\alpha}{\sim}$ is $n$-equivalent to $\chi_0$.

		Assume now that the claim holds for each $m<m^\prime\in\posnats$.  Let $m=m^\prime$ and suppose to the contrary that for every $F^\prime\subseteq K$ and every (nonempty) open interval $J\subseteq\faktor{\alpha}{\sim}$ there exists some $\chi_k\in F^\prime$ such that $0\leq k<m$ and $D_{\chi_k}(J)$ is not dense in $J$.  Note that, by assumption, there must exist a $\chi\in K$ such that $D_{\chi}(\faktor{\alpha}{\sim})$ is not dense in $\faktor{\alpha}{\sim}$.  Without loss of generality we may assume that $\chi=\chi_{m-1}$ and consequently there exists a (nonempty) open subset $J$ of $\faktor{\alpha}{\sim}$ such that $D_{\chi_{m-1}}(J)$ is not dense in $J$.  Note in particular that we can choose $J$ such that $\chi_{m-1}$ has no $n$-equivalent in $J$.

		By definition, there exists an $F^\prime\subseteq\set{\chi_0,\dotsc,\chi_{m-2}}$ which is an $n$-spectrum for the class $\setbuild{\chi^\prime\in\M}{\chi\nequiv{n}\gamma\text{ for some }\gamma\in J}$.  It then follows from the inductive hypothesis that there exists a maximal $F\subseteq F^\prime$ and some open interval $I=(A,B)\subseteq J$, where $A<B$, such that $D_{\chi}(I)$ is dense in $I$, for each $\chi\in F$.  Since $\chi_{m-1}\notin J\supseteq I$, it follows that $\chi_{m-1}$ is not dense in $I$ and, consequently, $F\subseteq K$ is maximal, as required.  This is the sought-after contradiction, thus establishing the claim.
	\end{proof}
	Note that $F$ is now an $n$-spectrum for the class
	\begin{equation}
		\setbuild{\chi\in\M}{\chi\nequiv{n}\gamma\text{ for some }\gamma\in I}.
	\end{equation}
	Consequently, since $D_{\chi}(I)$ is dense in $I$ for each $\chi\in F$, it follows that
	\begin{equation}
		\sigma(F)\nequiv{n}\sum I=\sum_{\gamma\in I}\gamma.
	\end{equation}
	Therefore, since $\sigma(F)\in\M$, it follows that $I$ has exactly one element.  This then contradicts the fact that $I$ is an open interval.  Hence, we may now conclude that $\faktor{\alpha}{\sim}$ has only one element, yielding the result.

\end{proof}

\begin{lem}\label{lem:relin}
	The set $\setbuild{(\alpha,\sigma)}{\alpha\in\M\text{ and }\alpha\models\sigma}$ is recursively enumerable.
\end{lem}
\begin{proof}
	The proof is near identical to that of Lemma \label{lem:rescat}.  We shall only highlight the additions necessary to prove the current result.

	It is necessary to add an additional clause, respectively, to the defintions of $L_\alpha$, $S_\alpha$ and $T_\alpha$.  If, for some $n\in\nats$, there exists $\alpha_0,\dotsc,\alpha_{n-1}\in\M\setminus\set{\zero}$ such that $\alpha=\sigma(\alpha_0,\dotsc,\alpha_{n-1})$ then $L_\alpha$ is obtained from $L$ by adding new unary relation symbols $r_0,\dotsc,r_{n-1}$ and a new binary relation symbol $\theta$.

	We choose $S_\alpha$ to consist of sentences expressing that ``$\theta$ is a congruence relation'' as well as, for each $i<n$, the sentences:
	\begin{itemize}
		\item	$\exists x r_i(x)$,
		\item	$\forall x\big(r_0(x)\vee\dotsb\vee r_{n-1}(x)\big)$,
		\item	for each $j<i$: $\neg\exists x\big(r_i(x)\wedge r_j(x)\big)$,
		\item	$\forall x\forall y\big(x<y\rightarrow\exists z(r_i(z)\wedge x<z \wedge z<y))\big)$,
		\item	$\forall x\big(r_i(x)\rightarrow\forall y(\theta(y,x)\rightarrow r_i(y))\big)$.
		\item	$\forall x\big(\exists y(\neg\theta(x,y)\wedge y<x)\wedge\exists y(\neg\theta(x,y)\wedge x<y)\big)$
	\end{itemize}

	Define the $L_\alpha$-theory
	\begin{equation}
		T_\alpha=S_\alpha\cup\bigcup_{i<n}\setbuild{\forall x\left(r_i(x)\rightarrow\sigma^{\theta(v,x)}\right)}{\alpha_i\models\sigma}
	\end{equation}
	and note that, by definition, $T_\alpha$ is recursively enumerable.  Hence, we need only show that $T_\alpha$ is complete.  To this end, consider any $\mathfrak{M},\mathfrak{N}\models T_\alpha$.

	Fix any $k\in\nats$, so that our task is reduced to showing that $\mathfrak{M}\nequiv{k}\mathfrak{N}$.  This is facilitated by the following description of a winning strategy for player $\Right$ in the game $\EF_k(\mathfrak{M},\mathfrak{N})$.

	Should it be the case that $n\leq 1$ then it follows that $\mathfrak{M}\equiv\mathfrak{N}$ so we may suppose that $n>1$.  Let $a_0\in M$ be the first move of player $\Left$.  By definition of $T_\alpha$, there exists a natural $i_0<n$ such that $\mathfrak{M}\models r_{i_0}(a_0)$.

	It then follows from the definition of $T_\alpha$ that there exists a countermove $b_0\in N$ for player $\Right$ such that $\mathfrak{N}\models r_{i_0}(b_0)$.  Any such $b_0$ will suffice and a similar tactic is used if $\Left$ instead chooses his first move in $\mathfrak{N}$.

	Suppose now that $\bar{a}\in M^{k-1}$ and $\bar{b}\in N^{k-1}$ and we are in the position $(\bar{a},\bar{b})$ in the game.  Without loss of generality we may assume $a_i<a_{i+1}$, for $i<k$, and need only consider the case where $\Left$ never repeats any of his moves.

	Suppose, during the $k$-th round, $\Left$ plays some $c\in M$.  There must then exist, by definition of $T_\alpha$, some $i_k<n$ such that $\mathfrak{M}\models r_{i_{k-1}}(c)$.  If $c>a_{k-1}$ then there must exists some $d\in N$ such that $d>\bar{b}$ and $\mathfrak{N}\models r_{i_{k-1}}(d)$.  A similar scenario is encountered if $c<a_0$: there exists a $d<\bar{b}$ in $\mathfrak{M}$ such that $\mathfrak{N}\models r_{i_{k-1}}(d)$.

	Now, consider the case where $a_j<c<a_{j+1}$, for some $j<k$.  If $b_{j+1}\leq b_j$ then $\Right$ may play any element but otherwise there exists a $d\in N$ such that $b_j<d<b_{j+1}$ and, for $i<n$,
	\begin{equation}
		\mathfrak{M}\models r_i(c)\quad\iff\quad\mathfrak{N}\models r_i(d).
	\end{equation}

	Should $\Left$ instead play a move $d\in N$ then $\Right$ plays as follows.  If $b_{j+1}\leq b_j$, for some $j<k$ then $\Right$ may play any move in $\mathfrak{M}$ but otherwise $\Right$ plays similarly to the previous case where $\Left$ chose a move in $\mathfrak{M}$.  The roles of the two structures are merely interchanged in this case.

	Define the set
	\begin{equation}
		R^\prime=\setbuild{(\alpha,\sigma)}{\alpha\in\M\text{ and }\alpha\models\sigma}.
	\end{equation}
	We now adapt the algorithm given in the proof of Lemma \ref{lem:rescat} with the goal of listing the elements of $R^\prime$.  This is accomplished by also taking shuffles into account.

	Assume the entries $(\alpha_i,\sigma_i)$, for $i=0,\dotsc,n_0-1$, have been listed.  Continue the list sequentially with entries $(\alpha_i,\sigma_i)$, for $i\geq n_0$, as follows:
	\begin{enumerate}
		\item 	list all pairs of the form $(\alpha_i,\sigma)$ such that $\sigma\in S_{\alpha_i}$ and $i<n_0$ and let $n_1$ denote the resulting number of entries in the list;
		\item 	list all pairs of the form $(\alpha_i, \sigma)$, for $i<n_1$, where $\sigma$ is the direct consequence of an inference rule from $\sigma_{i_0},\dotsc,\sigma_{i_{k-1}}$ and, for each $j<k$, $(\alpha_i,\sigma_{i_j})$ has already appeared in the list and let $n_2$ denote the resulting number of entries;
		\item	list all pairs of the form $(\alpha,\sigma)$ where, for $i,j<n_2$, $\alpha=\alpha_i+\alpha_j$ and $\sigma=\sigma_i^{r_0(v)}\wedge\sigma_j^{r_1(v)}$ and let $n_3$ denote the resulting number of entries in the list;
		\item	list all pairs of the form $(\alpha_i\cdot\omega,\sigma)$, for $i<n_3$, such that $\sigma=\forall w\sigma_i^{\theta(w,v)}$ and let $n_4$ denote the resulting number of entries in the list;
		\item	list all pairs of the form $(\alpha_i\cdot\dual{\omega},\sigma)$, for $i<n_4$, such that $\sigma=\forall w\sigma_i^{\theta(w,v)}$ and let $n_5$ denote the resulting number of entries in the list;
		\item	for each $k<n_5$ and nonzero linear orders $\alpha_{i_0},\dotsc,\alpha_{i_k}$, add $(\alpha,\sigma)$ to the list whenever
		      \begin{equation}
			      \alpha=\sigma(\alpha_{i_0},\dotsc,\alpha_{i_{k-1}})
		      \end{equation}
		      and, for some $j<k$,
		      \begin{equation}
			      \sigma=\forall x\left(r_{i_j}(x)\rightarrow\sigma_{i_j}^{\theta(v,x)}\right);
		      \end{equation}
		\item	redefine $n_0$ to denote the current length of the list;
		\item	repeat steps 1 to 7.
	\end{enumerate}

	An argument by means of induction (on $\alpha\in\M$) will now show that the procedure described above will list any desired member of $R^\prime$ after a finite amount of time.  Consequently, the same may be said of $R$.
\end{proof}

\begin{thm}
	The theory of linear orders is decidable.
\end{thm}
\begin{proof}
	Since the theory of linear orders is finitely axiomatisable, say by $T=\set{\axmlin}$, there exists a machine $M_0$ that lists all of its consequences.  This is accomplished, using inference rules, by mechanically generating all proofs from the theory $T$.

	From Lemma \ref{lem:relin}, there exists a machine $M_1$ that mechanically lists all pairs $(\alpha,\sigma)$ such that $\alpha\in\M$ and $\alpha\models\sigma$.  Now consider a machine $M$ that, given a sentence $\sigma$, alternately lists entries from each of the two lists until either $\sigma$ or a pair of the form $(\alpha,\neg\sigma)$ appears on the list.  In the former case $M$ prints the string ``True'' and in the latter it prints ``False''.

	It follows from Theorem \ref{thm:LLlin} and the definitions of $M_0$ and $M_1$ that $M$ always halts, irrespective of the choice of $\sigma$.  This is as required.
\end{proof}


\section{Model-theoretic aspects of the class \text{$\dense$}}

The class $\dense$ of dense linear orders without endpoints is relatively
well-behaved, from a model-theoretic perspective.  In this section we show that,
up to isomorphism, $\Th(\dense)$ has only a single countable model, namely the
rationals $\eta$.  This is a standard result known as Cantor's Theorem and it
establishes the property it establishes is known as $\aleph_0$-categoricity of
$\Th(\dense)$.  Further, we take a brief look at some other, equally
interesting, model-theoretic properties exhibited by this theory: quantifier
elimination, model-completeness and posession of a so-called prime model.

With $\alpha$ and $\beta$ both countable, consider a modification
$G=G(\alpha,\beta)$ of the game $\EF_{\omega}(\alpha,\beta)$, defined as
follows.  A play $(\bar{c},\bar{d})$ in the game $G$ is winning for $\Left$ iff
the following holds:
\begin{enumerate}[noitemsep]
	\item The player $\Left$ never plays an element from the same structure on
	      two consecutive turns.
	\item At least one of the following is satisfied:
	      \begin{enumerate}
		      \item the play $(\bar{c},\bar{d})$ is winning for $\Left$ in
		            $\EF_\omega(\alpha,\beta)$,
		      \item there exists a member of $\alpha$ not occuring in $\bar{c}$ or
		            there exists a member of $\beta$ not occuring in $\bar{d}$.
	      \end{enumerate}
\end{enumerate}

Note, for $\alpha$ and $\beta$ both countable, it follows by definition
that $\Right$ has a winning strategy for $G$ iff there exists an isomorphism
$f\colon\alpha\to\beta$.  The forward implication can be shown by considering
any play of the game in which $\Left$ plays optimally by alternating between
$\alpha$ and $\beta$.

Conversely, any such isomorphism $f$ will determine a winning strategy for
$\Right$ in the following manner.  A move $a$ in $\alpha$ by $\Left$ is
countered by playing $f(a)$ while a move $b$ in $\beta$ is countered by
$\inv{f}(b)$.

\begin{thm}[Cantor's Theorem]\label{thm:cantor}
	If $\alpha,\beta\in\dense$ is countable then there exists an isomorphism
	$f\colon\alpha\to\beta$.
\end{thm}
\begin{proof}
	Let $G=G(a,b)$ be the game described previously and fix enumerations
	$\bar{a}$ and $\bar{b}$ of $\alpha$ and $\beta$, respectively.  Note that
	our task is achieved if we can exhibit a winning strategy $\sigma$ for
	$\Right$ in $G$.

	We now describe the necessary $\sigma$ to conclude the proof in a recursive
	fashion.  Any initial move by $\Left$ can be countered by any choice of
	element in the opposite structure.

	Suppose now that $n-1$ rounds have been played, for some $n\geq 2$.  As the
	cases are approached similarly, we may assume $\Left$ played his initial
	move in $\alpha$.  It is also reasonable that $\Left$ plays optimally in the
	sense that he never plays from the same structure in two of his consecutive
	turns.

	Now, if $n$ is odd then $\Left$'s last move was some $a_{k_{n-1}}\in\alpha$.
	The proper response for $\Right$ in this scenario would be to choose the
	first $b_{j_{n-1}}$ occuring in $\bar{b}$ such that
	\begin{equation}
		(\alpha,a_{0},\dotsc,a_{k_{n-1}})\nequiv{0}(\beta,b_{0},\dotsc,b_{j_{n-1}}),
	\end{equation}
	the existence of which is guaranteed by the density of $\beta$.  In other
	words, the (simple) expansions above should satisfy the same atomic
	sentences.

	If, instead, $n$ is even then $\Left$ has just played some $b_{k_{n-1}}$ so
	that the appropriate response is similarly determined.  The only difference
	being the reversal of the roles of the structures: $\Left$ chose from $\beta$
	and $\Right$ will respond with an element from $\alpha$.
\end{proof}

\begin{prp}\label{prp:xcantor}
	If $\alpha\in\dense$ is countable and, for some $n\in\nats$, we fix any
	$\bar{a}\in\domain{n}\alpha$ and any $\bar{b}\in\rats^n$ then there exists an
	isomorphism $f\colon\alpha\to\eta$ such that $f[\bar{a}]=\bar{b}$.
\end{prp}
\begin{proof}
	We prooceed by means of induction, nothing that the case $n=0$ is Cantor's
	Theorem (Theorem \ref{thm:cantor}).  Suppose now the result holds for each
	$k\leq n$ for some arbtrary $n\in\nats$ and note that we may assume
	$\bar{a}=(a_0,\dotsc,a_n)$ and $\bar{b}=(b_0,\dotsc,a_n)$ such that $a_i<a_j$
	and $b_i<b_j$, whenever $i<j<\omega$.  It now follows from the inductive
	hypothesis that there exists isomoprhisms $f_0\colon\alpha^{<a_0}\to\eta^{<b_0}$
	and $f_1\colon(\alpha^{>a_0},a_1,\dotsc,a_n)\to(\beta^{>a_0},b_1,\dotsc,b_n)$.
	By identifying $f_0$ and $f_1$ with their (respective) underlying sets, we may
	define the new map $f\colon\alpha\to\beta$ such that
	$f=f_0\cup\set{(a_0,b_0)}\cup f_1$.  By definition, we have thusly obtained an
	order-preserving map $f\colon(\alpha,\bar{a})\to(\beta,\bar{b})$.  It is not
	difficult to see, simply from its definition, that $f$ is in fact an isomorphism
	$f\colon\alpha\to\beta$ satisfying $f[\bar{a}]=\bar{b}$, as required.
\end{proof}

As a consequence of the previous proposition we get the following:
\begin{prp}
	The linear order $\eta$ is saturated.
\end{prp}
\begin{proof}
	Choose any finite $A\subseteq\rats$.  Suppose that $\Phi(x)\subseteq L_1(A)$
	is an $1$-type of $\eta$ over $A$.  By definition there exists a
	$\alpha\succcurlyeq\eta$ such that $\Phi$ is realised by some $b\in\alpha$ i.e.\
	$\alpha\models\Phi(b)$.  By the Downwards L\"owhenheim-Skolem Theorem, however,
	we may assume that $\alpha$ is countable.

	Note we may set $A=\set{a_0,\dotsc,a_{k-1}}$ and thus, from Proposition
	\ref{prp:xcantor}, if $\bar{a}=(a_0,\dotsc,a_{k-1})$ then there must exist an
	isomorphism $f\colon(\eta,\bar{a})\to(\alpha,\bar{a})$.  Therefore, it must
	follow that $\eta\models\varphi(f(b))$, for each $\varphi\in\Phi(x)$, and thus
	$\Phi(x)$ is realised by the element $f(b)\in\rats$.
\end{proof}


\begin{thm}
	The linear order $\eta$ is a prime model of the theory $\Th(\dense)$.
\end{thm}
\begin{proof}
	Suppose $\alpha\models\Th(\dense)$ then $\alpha$ must be infinite and thus, from the downward L\"owenheim-Skolem Theorem, there exists a countable $\alpha_0\preccurlyeq\alpha$.  From Cantor's theorem it follows that $\alpha_0\cong\eta$ and, therefore, there must exist an elementary embedding of $\eta$ into $\alpha$.
\end{proof}

\begin{thm}
	The theory $T=\Th(\dense)$ has quantifier elimination.
\end{thm}
\begin{proof}
	By Lemma \ref{lem:qelim}, it is enough to show that $T$ is substructure complete.  Suppose that $\alpha\models T$ and choose any $\alpha_0\subseteq\alpha$.  Define $A=\domain{}\alpha_0$ and suppose that $\mathfrak{M},\mathfrak{N}\models T\cup\diag{\alpha_0}$ are $L(A)$-structures.  Note that we may assume, without loss of generality, that $A\subseteq M,N$ such that $\mathfrak{M}=(\beta,A)$ and $\mathfrak{N}=(\gamma,A)$ for linear orders $\beta,\gamma\models T$.  We are now required to show that
	\begin{equation}
		\mathfrak{M}\equiv\mathfrak{N}.
	\end{equation}
	However, it is enough to simply prove that
	\begin{equation}
		(\beta,B)\equiv(\gamma,B),\quad\text{for every finite }B\subseteq A.
	\end{equation}

	Suppose that $B\subseteq A$ is finite then by the downwards
	L\"owenheim-Skolem Theorem there exists countable $\beta_0\elmsub\beta$ and
	$\gamma_0\elmsub\gamma$ such that $B\subseteq\beta_0,\gamma_0$.  We may assume
	that the elements of $B$, for some $n\in\nats$, are precisely
	$b_0<b_1<\dotsb<b_{n-1}$.  It now follows from Cantor's Theorem that each of the
	intervals $(b_0,b_1)_{\beta_0},\dotsc,(b_{n-2},b_{n-1})_{\beta_0}$ as well as
	$\beta_0^{<b_0}$ and $\beta_0^{>b_{n-1}}$ are isomorphic to $\eta$.  The same is
	true of the corresponding intervals in $\gamma_0$.  This then gives rise to
	isomorphisms $f_0,\dotsc,f_n$ (whose domains we will choose to be subsets of
	$\domain{}\beta_0$) between the corresponding open intervals of $\beta_0$ and
	$\gamma_0$ such that each member of the range of $f_i$ is less (in $\gamma_0$)
	than every member member of the range of $f_j$ whenever $i<j\leq n$.  One can
	then define (in an obvious manner) an isomorphism $f\colon\beta_0\to\gamma_0$
	which is an extension of $f_i$ to all of $\beta_0$, for each $i\leq n$, such
	that $f(b_j)=b_j$ whenever $j<n$.  Therefore, since no element of $B$ is an
	endpoint of either $\beta_0$ or $\gamma_0$, it follows that
	$(\beta_0,B)\cong(\gamma_0,B)$ and thus $(\beta_0,B)\equiv(\gamma_0,B)$.  The
	proof is then concluded by recalling that $\beta_0\elmsub\beta$ and
	$\gamma_0\elmsub\gamma$.
\end{proof}

\begin{cor}
	The theory $T=\Th(\dense)$ is model-complete.
\end{cor}
\begin{proof}
	Immediate from the previous theorem and Corollary \ref{cor:qemc}.
\end{proof}

We've already seen that $\eta$ is saturated and can thus be thought of from an intuitive standpoint as being very ``thick'' among linear orders.  Contrary to what one might expect, though, the next result shows that $\eta$ is simultaneously also an atomic structure and can therefore be thought of as being very ``thin''!

\begin{prp}
	The linear order $\eta$ is atomic.
\end{prp}
\begin{proof}
	Let $T=\Th(\dense)$ then the result follows from the fact that $T$ admits quantifier elimination and the fact that there are only finitely many inequivalent formulas in finitely many variables in a finitary language.
\end{proof}
